---
title: Programmatically extract TIOBE Index Ratings
author: AbdulMajedRaja RS
date: '2019-08-03'
slug: programmatically-extract-tiobe-index-ratings
categories: 
  - How-To
tags:
  - r-bloggers
  - tiobe
  - programming
description: 'using tiobeindexr R package'
topics: []
---

**TIOBE Index** is an index (ranking) that claims to represent the popularity of programming languages. [Yihui](https://yihui.name/en/about/) (The creator of `blogdown` package), recently wrote a blogpost titled "[On TIOBE Index and the era of decision fatigue](https://yihui.name/en/2019/06/tiobe-index/)" and I strongly recommend you to go through that before continuing with this post. 

So the **Disclaimer** goes like this: *This post/author doesn't believe that TIOBE Index is a fair way to measure/present popularity of programming languages and this is writtet just to teach you how to extract/get TIOBE Index programmatically using the R package `tiobeindexr` *

### tiobeindexr - Intro, Installation & Loading

`tiobeindexr` is an R package to extract TIOBE Index of the given month. 

`tiobeindexr` is available on CRAN so you can install like below:

```{r eval=FALSE}
install.packages("tiobeindexr")
```

Once installed, it can be loaded like any other R package:

```{r}
library(tiobeindexr)
```

When `tiobeindexr` is loaded for the first time in the given session, it downloads the required data from the internet. 

### Extract top 20 programming languages of the month 

TIOBE Index publishes the rank of programming languages every month (monthly-refresh). We can use the function `top_20()` to extract the top 20 programming languages of the month (that TIOBE has published)

```{r}
top_20()
```

### Visualising Top Changes of TIOBE Index This Month vs Previous Month

As you can see in the output of the previous section, `top_20()` also gives us the % change MoM (Month-over-Month) which we can use to see the top changes. 

For simplicity, We'll load the entire `tidyverse` package and use `ggplot2`'s bar-plot to visualize the changes. 

```{r warning=FALSE, message=FALSE}
library(tidyverse)

top_20() %>% 
  mutate(Change = as.numeric(gsub('%','',Change))) %>% 
  ggplot(aes(x = reorder(`Programming Language`,Change), y = Change, 
             fill = `Programming Language`,
             label = paste0(Change, "%"))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  geom_text(nudge_x = 0.1) +
  xlab('Programming Language') +
  ggtitle('Programming Languages Change this Month') 
```

### Summary

Hence, we learnt how to use `tiobeindexr` to programmatically download `TIOBE Index` and visualize insights from it. And, alongisde we also learnt that TIOBE Index in fact isn't a fair represenation of the popularity of programming languages

### References 

+ [tiobeindexr - Github](https://github.com/amrrs/tiobeindexr)
+ [TIOBE Index - Official Page](https://www.tiobe.com/tiobe-index/)
