---
title: Interactive Visualization in R with apexcharter
author: AbdulMajedRaja RS
date: '2019-08-04'
slug: interactive-visualization-in-r-with-apexcharter
categories:
  - Data Visualization
  - How-To
tags:
  - r-bloggers
  - R Packages
  - Data Visualization
  - Interactive Visualization
description: 'Build interactive charts and visualizations'
topics: []
---

Interactive Visualizations are powerful these days because those are all made for web. Web - simply a combination of `html`,`css` and `javascript` which build interactive visualizations. Thus, paving way for a lot of javascript charting libraries like `highcharts.js`, `apexcharts.js`. 

Thanks to `htmlwidgets` of R, many R developers have started porting those javascript charting libraries to R and [dreamRs](https://www.dreamrs.fr/) is one of such leading Developer groups working on the intersection R + Web. In this post, We'll learn how to use the R package `apexcharter` which is developed by **dreamRs - Victor Perrier and Team** to make beautiful interactive visualizations that are based on `apexcharts.js`

### apexcharter - Intro, Installation & Loading

[`apexchart.js`](https://apexcharts.com/) is a modern JavaScript charting library to build interactive charts and visualizations with simple API. `apexcharter` is built as a htmlwidget (R Package) for `apexchart.js` and the API design is inspired by `highcharter`. `apexcharter` requires RStudio >= 1.2 to properly display charts.

Install the stable version from CRAN with:

```r
install.packages("apexcharter")
```

Or install the development version from [GitHub](https://github.com/dreamRs/apexcharter/) with:

``` r
# install.packages("devtools")
devtools::install_github("dreamRs/apexcharter")
```

Once successfully installed, `apexcharter` can be loaded using 
``` r
library(apexcharter)
```

### Simple Example

The main function of `apexcharter` is the `apex()` function whose first argument is `data`. Thus, enabling the support of pipe `%>% ` operator. The second argument is `mapping` - aesthetics (x & y) and the third one is `type` of the chart - which takes multiple values like `scatter`, `bar`, `line` and much more.

Let's take R's in-built `mtcars` dataset and draw a simple bar chart. 

```{r warning=FALSE, message=FALSE}

library(apexcharter)
library(tidyverse)

mtcars %>% 
  count(cyl) %>% 
  apex(type = "bar",
       mapping = aes(x = "cyl", y = n))

```

Now, that's a beautiful interctive chart. Let's go ahead and see a few more examples of something bigger than a simple bar chart. 

### Building Interactive Heatmap / Correlation Plot

Let's try to visualize a Heatmap (of Correlation Plot) of numeric columns of `mtcars` dataset. To do that, we've to first select the numeric columns which we'll do with `select_if(is.numeric)` and then we've to build the correlation matrix which the base-R function `cor()` does it smoothly. 

Now that we've got a `matrix` let's convert it to a `data.frame` and for us to draw a heatmap - we need 3 things primarily:

+ x-axis - categorical
+ y-axis  - categorical
+ fill value - continuous

So, we'll convert the rownames of the resultant dataframe to a column and then [convert the wide format data into long format](https://www.programmingwithr.com/how-to-reshape-a-dataframe-from-wide-to-long-or-long-to-wide-format/) using `pivot_longer()`. 

At this point our data is in the desirable format for a heatmap. Simply for aesthetics improvement, let's round off the correlation values. 

Finally, we'll use our `apex()` function with `type = 'heatmap'` that gives us a color-filled heatmap (that's also interactive).


```{r warning=FALSE, message=FALSE}
library(apexcharter)
library(tidyverse)

mtcars %>% 
  select_if(is.numeric) %>% 
  cor() %>% 
  as.data.frame() %>% 
  rownames_to_column("col") %>% 
  pivot_longer(cols = -col, names_to = "type") %>% 
  mutate(value = round(value,2)) %>% 
  apex(type = "heatmap",
       mapping = aes(x = col, y = type, fill = value)) 

```

### Building Interactive Time-Series (Line) Graph

If there's a plot where Interactive Charts are incredibly valuable, I think it's Time-Series Graph where labelling on traditional (static) chart would sometimes make the chart clunky and less readable. 

Let's build an Interactive Time-series plot with the `apexcharter` library. As you can see below, all it takes is a dataframe with a column denoting the `time` field and another column with the actual `value` for that time. 


```{r}

library(apexcharter)

df <- data.frame(Y=as.matrix(EuStockMarkets), date=time(EuStockMarkets))

df %>% 
  apex(type = "line",
       mapping = aes(x = date, y = Y.DAX)) 

```

### Summary

Thus, We learnt how to build interactive charts using `apexcharter` that follows a very minimal API similar to `highcharter`. 

### References

+ [apexcharter : dreamRs - Github](https://github.com/dreamRs/apexcharter)
+ [apexcharts.com](https://apexcharts.com/)
+ [How to reshape a dataframe from wide to long or long to wide format using tidyr - pivot_longer() pivot_wider()](https://www.programmingwithr.com/how-to-reshape-a-dataframe-from-wide-to-long-or-long-to-wide-format/)



